\begin{Problem}[Density Matrix of a Single Qubit]
For a given ensemble of qubits the observables
\[
A=
\begin{pmatrix}
3 & 0\\
0 & 1
\end{pmatrix},
\qquad
B=
\begin{pmatrix}
1 & 1\\
1 & -1
\end{pmatrix},
\qquad
C=
\begin{pmatrix}
0 & 2i\\
-2i & 0
\end{pmatrix},
\]
have been measured many times, giving the expectation values
\[
\langle A\rangle = 2,
\qquad
\langle B\rangle = \frac12,
\qquad
\langle C\rangle = 0.
\]

\begin{enumerate}
\item[(a)] Determine the density matrix of the ensemble.
\item[(b)] Is the ensemble pure or mixed?
\item[(c)] If we measure $\sigma^z$, what is the probability to get $+1$?
\item[(d)] Determine the expectation values
\[
\langle \sigma^x\rangle,\qquad
\langle \sigma^y\rangle,\qquad
\langle \sigma^z\rangle.
\]
\end{enumerate}
\end{Problem}
\begin{proof}
	\begin{parts}
	\item Given a density matrix $\rho$, the expectation value of an operator $\mathbf{A}$ is given by $\langle \mathbf{A}\rangle = \Tr(\rho \mathbf{A})$. We implement the trace 1 and hermiticity conditions by defining our density matrix using
		\[
			\rho = \begin{pmatrix} a & b \\ b^* & 1-a \end{pmatrix} 
		.\] 
		Substituting this in, we find
		\begin{align*}
			\langle \mathbf{A}\rangle &= 1+2a = 2 \\
			\langle \mathbf{B}\rangle &= -1+2a+b+b^* = \frac{1}{2} \\
			\langle \mathbf{C}\rangle &= -2ib + 2ib^*=0 
		\end{align*}
		From the third equation, we see that $b=b^*$, i.e. $b$ is real. From the first equation, we can see easily that $a = \frac{1}{2}$. Substituting this in to the second equation, we find that $2b = \frac{1}{2}$, or $b = \frac{1}{4}$.
	\item We do this using the trace criterion: We have
		\[
			\rho^2 = \begin{pmatrix} \frac{5}{16} & \frac{1}{4} \\ \frac{1}{4} & \frac{5}{16} \end{pmatrix} 
		.\] 
		Thus, the trace is
		\[
		\Tr(\rho^2) = \frac{10}{16}\neq 1
		.\] 
		The ensemble is mixed.
	\item ..
	\item We compute
		\[
			\rho \sigma^z = \begin{pmatrix} \frac{1}{2} & -\frac{1}{4} \\ \frac{1}{4} & -\frac{1}{2} \end{pmatrix} 
		.\] 
		Thus, we have
		\[
		\langle \rho\sigma^z\rangle = 0
		.\] 
		Similarly, we have
		\[
			\rho\sigma^x = \begin{pmatrix} \frac{1}{4} & \frac{1}{2} \\ \frac{1}{2} & \frac{1}{4} \end{pmatrix} 
		\]
		with
		\[
			\langle \rho \sigma^x\rangle = \frac{1}{2}
		\]
		and
		\[
			\rho \sigma^y = \begin{pmatrix} \frac{i}{4} & -\frac{i}{2} \\ \frac{i}{2} & -\frac{i}{4} \end{pmatrix} 
		\]
		with
		\[
		\langle \rho\sigma^y \rangle = 0
		.\] 
	\end{parts}
\end{proof}

\begin{Problem}[Criterion for the Equivalence of Ensembles]
Consider a statistical ensemble $\{|\psi_i\rangle,p_i\}$ with the associated
density matrix
\[
\rho=\sum_{i=1}^{N} p_i |\psi_i\rangle\langle\psi_i|.
\]

Here $N$ is arbitrary (smaller or larger than the dimension of the Hilbert
space) and the vectors $|\psi_i\rangle$ are pairwise different and normalized
but not necessarily orthogonal. Moreover, let
\[
\rho=\sum_{k=1}^{K}\lambda_k |k\rangle\langle k|
\]
be the spectral decomposition of $\rho$ with orthonormal eigenvectors $|k\rangle$
and eigenvalues $\lambda_k$.

The aim of this exercise is to work out an equivalence theorem for different
ensembles with the same density matrix.

\begin{enumerate}
\item[(a)] Show that every vector $|\psi_i\rangle$ can be written as a linear
combination of the eigenvectors $|k\rangle$, i.e.
\[
|\psi_i\rangle\in \mathrm{span}\{|k\rangle\}.
\]

\item[(b)] Use (a) to prove that every $|\psi_i\rangle$ can be written as
\[
\sqrt{p_i}\,|\psi_i\rangle
=
\sum_{k=1}^{K}
c_{ik}\sqrt{\lambda_k}\,|k\rangle,
\]
with coefficients $c_{ik}$ obeying
\[
\sum_{i=1}^{N} c_{ik}c_{il}^{*}
=
\delta_{kl},
\qquad
k,l=1,\dots,K.
\]

\item[(c)] Because of (a) we know that $N\ge K$. For $N=K$ the result of (b)
would imply that the coefficient matrix $c_{ik}$ is unitary. Show that for
$N>K$ the rectangular matrix $c_{ik}$ can be extended to a square unitary
matrix such that the relation in (b) still holds.

\item[(d)] So far we have proven that if
\[
\sum_{i=1}^{N} p_i |\psi_i\rangle\langle\psi_i|
=
\sum_{k=1}^{K}\lambda_k |k\rangle\langle k|,
\]
then there exists a unitary $N\times N$ matrix $c_{ik}$ such that
\[
\sqrt{p_i}\,|\psi_i\rangle
=
\sum_{k=1}^{N}
c_{ik}\sqrt{\lambda_k}\,|k\rangle.
\]
Prove the converse statement.

\item[(e)] Use (a)--(d) to show that
\[
p_i=\sum_{k=1}^{N} T_{ik}\lambda_k,
\]
(or in vector form $\vec p=T\vec\lambda$),
where $T$ is a doubly-stochastic matrix, i.e. all matrix entries are real
and nonnegative and
\[
\sum_{i=1}^{N}T_{ik}
=
\sum_{k=1}^{N}T_{ik}
=
1.
\]

\item[(f)] Use e.g. Mathematica to work out the explicit example
\[
p_1=p_2=\frac12,
\qquad
|\psi_1\rangle
=
\frac1{\sqrt2}
\begin{pmatrix}
1\\
1
\end{pmatrix},
\qquad
|\psi_2\rangle
=
\begin{pmatrix}
0\\
1
\end{pmatrix},
\]
with $N=K=2$, and determine
$\lambda_{1,2}$, $|\phi_{1,2}\rangle$, $c_{ij}$, and $T_{ij}$.
Check the properties derived in (e).
\end{enumerate}
\end{Problem}
\begin{proof}
	\textbf{Brief note on interpretation:} In the sum
	\[
		\rho = \sum_{k=1}^K \lambda_k \dyad{k}{k}
	\]
	the sum only contains the eigenvectors with \textbf{nonzero} eigenvalue. This is in contrast to the standard way the spectral theorem is written including the kernel. Otherwise, (a) would follow trivially from the spectral theorem, because the vectors $\ket{k}$ would span the space, and the opening statement in (c) would be false, because $K$ would be the dimension of the Hilbert space, and it is not necessarily true that $N\ge \dim \mathcal{H}$.
	\begin{parts}
	\item As from the comment, in this problem we basically seek to show that $\ket{\psi_i}$ is spanned by the eigenvectors with nonzero eigenvalue. However, we also know that the eigenvectors with nonzero eigenvalue span the kernel. Thus, we decompose
		\[
			\mathcal{H }= \ker \rho \oplus \text{im } \rho = \ker\rho \oplus (\ker\rho)^\perp
		.\] 
		We seek to show that $\ket{\psi_i} \in \text{im }\rho$. Since this is an orthogonal decomposition, it suffices to show that $\braket{\phi}{\psi_i}=0$ for all $\ket{\phi}\in \ker \rho$, i.e. $\ket{\psi_i}$ is orthogonal to $\ker\rho$.

		Consider $\ket{\phi}\in \ker\rho$. Then we have
		\[
			0 = \mel{\phi}{\rho}{\phi} = \sum_{i=1}^{N} p_i \braket{\phi}{\psi_i}\braket{\psi_i}{\phi} = \sum_{i=1}^N p_i |\braket{\phi}{\psi_i}|^2
		.\] 
		Since the terms in parentheses are all positive, it follows that $\braket{\phi}{\psi_i}=0$ for all $\ket{\psi_i}$. Since $\ket{\phi}$ was arbitrary, the conclusion follows.
	\item Since $\ket{\psi_i}$ is in the span of $\ket{k}$, by definition
		\[
			\ket{\psi_i} = \sum_{k=1}^n a_{ik}\ket{k}
		.\] 
		By redefining the coefficients $a_{ik} = c_{ik} \sqrt{\lambda_k}  / \sqrt{p_i} $, we can write
		\[
			\sqrt{p_i} \ket{\psi_i}  = \sum_{k=1}^K c_{ik} \sqrt{\lambda_k} \ket{k}
		.\] 
		Note that we require that $p_i \neq 0$, but this is given. Thus we need to solve the second part. First, we take the conjugate to obtain
		\[
			\sqrt{p_i} \bra{\psi_i} = \sum_{k=1}^K c_{ik}^* \sqrt{\lambda_k} \bra{k}
		.\] 
		We substitute these in the expression from $\rho$ to get
		\begin{align*}
			\rho &= \sum_{i=1}^N p_i \dyad{\psi_i}{\psi_i} \\
			     &= \sum_{i=1}^N p_i \left[ \frac{1}{\sqrt{p_i} }\sum_{k=1}^K c_{ik}\sqrt{\lambda_k} \ket{k} \right]\left[ \frac{1}{\sqrt{p_i} }\sum_{l=1}^K c_{il}^* \sqrt{\lambda_l}\bra{l}  \right]   \\
			     &= \sum_{i=1}^N\sum_{k=1}^K \sum_{l=1}^K c_{ik}c_{il}^* \sqrt{\lambda_k\lambda_l} \dyad{k}{l} 
		\end{align*}
		Now, since the $\ket{k}$ are orthogonal by the spectral theorem, we are allowed to compare coefficients. The off diagonal terms must vanish and we have
		\[
			\sum_{i=1}^N c_{ik}c_{il}^* = \delta_{kl}
		\]
		to obtain the density matrix.
	\item Note on the indices in $c_{ik}$: $i$ goes from 1 to $N$ and $k$ goes from 1 to $K<N$. Thus, for each $k$, we have a vector in $\C^N$. The $c$s are orthonormal for different $i$. By the basis extension theorem, we can extend this basis to an orthonormal basis of $\C^N$, i.e. pick $c_{ik}$ for $k>K$ such that $c_{ik}$s are orthonormal.
	\item We simply substitute the second equation into the first
		\begin{align*}
			\sum_{i=1}^N p_i \dyad{\psi_i}{\psi_i} &= \sum_{i=1}^N (\sqrt{p_i} \ket{\psi_i})(\bra{\psi_i} \sqrt{p_i} ) \\
							       &= \sum_{i=1}^N \left[ \sum_{k=1}^N c_{ik}\sqrt{\lambda_k} \ket{k} \right]\left[ \sum_{l=1}^N c_{il}^* \sqrt{\lambda_l} \bra{l} \right]   \\
							       &= \sum_{k=1}^N \sum_{l=1}^N \underbrace{\left[ \sum_{i=1}^N c_{ik}c_{il}^* \right]}_{\delta_{kl}} \sqrt{\lambda_k\lambda_l} \dyad{k}{l} \\
							       &= \sum_{k=1}^N \lambda_k \dyad{k}{k} 
		\end{align*}
	\item We take norms of the equation
		\[
			\sqrt{p_i} \ket{\psi_i} = \sum_{k=1}^n c_{ik}\sqrt{\lambda_k}\ket{k} 
		\]
		to get
		\[
			p_i = \sum_{k=1}^N |c_{ik}|^2 \lambda_k
		.\] 
		By inspection, we find
		\[
			T_{ik} = |c_{ik}|^2
		.\] 
		Clearly, all matrix entries are real and nonnegative. Now, we evaluate the sums
		\begin{align*}
			\sum_{k=1}^N T_{ik} &= \sum_{k=1}^N |c_{ik}|^2 \\
					    &= \sum_{k=1}^N c_{ik}c_{ik}^* \\
					    &= \sum_{k=1}^N c^\dagger_{ki}c_{ik} \\
					    &= \delta_{kk} = 1 
		\end{align*}
		The other sum follows basically identically:
		\begin{align*}
			\sum_{i=1}^N T_{ik} &= \sum_{i=1}^N |c_{ik}|^2 \\
					    &= \sum_{i=1}^N c_{ik}c_{ik}^* \\
					    &= \sum_{i=1}^N c^\dagger_{ki}c_{ik} \\
					    &= \delta_{kk}=1 
		\end{align*}
	\item We compute these in Mathematica. First, we directly compute
		\begin{mmaCell}{Input}
p1 = 1/2;
\end{mmaCell}
\begin{mmaCell}{Input}
p2 = 1/2;
\end{mmaCell}
\begin{mmaCell}{Input}
\mmaUnd{\(\psi1\)} = \{1, 1\}/\mmaSqrt{2};
\end{mmaCell}
\begin{mmaCell}{Input}
\mmaUnd{\(\psi2\)} = \{0, 1\};
\end{mmaCell}
\begin{mmaCell}{Input}
\mmaUnd{\(\rho\)} = \mmaUnd{p1} KroneckerProduct[\mmaUnd{\(\psi1\)}, \mmaUnd{\(\psi1\)}] + \mmaUnd{p2} KroneckerProduct[\mmaUnd{\(\psi2\)}, \mmaUnd{\(\psi2\)}]

		\end{mmaCell}
		\begin{mmaCell}{Output}
\{\{1/4, 1/4\}, \{1/4, 3/4\}\}
		\end{mmaCell}
to find
		\begin{align*}
			\rho &= p_1 \dyad{\psi_1}{\psi_1} + p_2 \dyad{\psi_2}{\psi_2} \\
			     &= \begin{pmatrix} \frac{1}{4} & \frac{1}{4}\\ \frac{1}{4 }& \frac{3}{4} \end{pmatrix} 
		\end{align*}
		Then, we compute the eigenvalues and eigenvectors using
\begin{mmaCell}{Input}
\{\mmaUnd{eigvals}, \mmaUnd{eigvecs}\} = Eigensystem[\mmaUnd{\(\rho\)}]
\end{mmaCell}
\begin{mmaCell}{Output}
\{\{1/4 (2 + Sqrt[2]), 1/4 (2 - Sqrt[2])\}, \{\{-1 + Sqrt[2], 1\}, \{-1 - Sqrt[2], 1\}\}\} 
\end{mmaCell}
		leading to eigenvalues
		\[
			\lambda_{\pm} = \frac{1}{4}(2\pm \sqrt{2} )
		\]
		and associated eigenvectors
		\[
			\ket{\phi_{\pm}} = \begin{pmatrix} -1 \mp \sqrt{2}  \\ 1 \end{pmatrix} 
		.\] 
		Although these vectors are orthogonal, they are not normalised. Thus, we divide through
		\begin{mmaCell}{Input}
eivecsnorm = #/Norm[#] & /@ eigvecs
		\end{mmaCell}
\begin{mmaCell}{Output}
\{\{(-1 + Sqrt[2])/Sqrt[1 + (-1 + Sqrt[2])^2], 1/Sqrt[1 + (-1 + Sqrt[2])^2]\}, \{(-1 - Sqrt[2])/Sqrt[1 + (1 + Sqrt[2])^2], 1/Sqrt[1 + (1 + Sqrt[2])^2]\}\}
\end{mmaCell}


		To compute the $c_{ik}$, we note that the $\ket{k}$ form an orthonormal basis by the spectral theorem. Thus, we have
		\[
			\sqrt{p_i}\braket{l}{\psi_i} = c_{il}\sqrt{\lambda_l} 
		.\] 
		Thus, we compute
		\[
			c_{ik} = \frac{\sqrt{p_i} \braket{k}{\psi_i}}{\sqrt{\lambda_k} }
		.\] 
		Defining $\ket{\phi_1} = \ket{\phi_+} = \ket{1}$ and $\ket{\phi_2}=\ket{\phi_-}=\ket{2}$, we thus evaluate these using
		\begin{mmaCell}{Input}
c = FullSimplify[Table[(Sqrt[p[[i]]] Conjugate[eivecsnorm[[k]]] . \mmaUnd{\(\psi\)}[[i]])/Sqrt[eigvals[[k]]], \{i, 1, 2\}, \{k, 1, 2\}]]
		\end{mmaCell}
		\begin{mmaCell}{Output}
\{\{1/Sqrt[2], - 1/Sqrt[2]\}, \{1/Sqrt[2], 1/Sqrt[2]\}\}
		\end{mmaCell}
The $T$ matrix then follows easily, from $T = |c|^2$ elementwise as
\[
	T = \frac{1}{2}\begin{pmatrix} 1 & 1 \\ 1 & 1 \end{pmatrix} 
.\] 
	\end{parts}
\end{proof}
