\documentclass[prb,12pt]{revtex4-2}

\usepackage{amsmath, amssymb,physics,amsfonts,amsthm}
\usepackage{enumitem}
\usepackage[most]{tcolorbox}
\usepackage{cancel}
\usepackage{booktabs}
\usepackage{tikz}
\usepackage{pgf}
\usepackage{xurl}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage[normalem]{ulem}
\usepackage{transparent}
\usepackage{float}
\usepackage{multirow}
\usepackage{subcaption}
\usepackage{mathtools}
\usepackage{thmtools}
\usepackage{thm-restate}
\usepackage[framemethod=TikZ]{mdframed}
\mdfsetup{skipabove=1em,skipbelow=0em, innertopmargin=12pt, innerbottommargin=8pt}
\declaretheoremstyle[headfont=\bfseries\sffamily, bodyfont=\normalfont, mdframed={  } ]{thmbox}
%put nobreak in mdframed
\declaretheorem[style=thmbox, name=Theorem]{Theorem}
\declaretheorem[sibling=Theorem, style=thmbox, name=Definition]{Definition}
\declaretheorem[sibling=Theorem, style=thmbox, name=Corollary]{Corollary}
\declaretheorem[sibling=Theorem, style=thmbox, name=Example]{Example}
\declaretheorem[sibling=Theorem, style=thmbox, name=Proposition]{Proposition}
\declaretheorem[sibling=Theorem, style=thmbox, name=Lemma]{Lemma}
%\newtheorem{Theorem}{Theorem}
%\numberwithin{Theorem}{chapter}
%\newtheorem{Proposition}{Proposition}
%\newtheorem{Lemma}[Theorem]{Lemma}
%\newtheorem{Corollary}[Theorem]{Corollary}
%\newtheorem{Example}[Theorem]{Example}
%\theoremstyle{definition}
\theoremstyle{definition}
\newtheorem{Remark}[Theorem]{Remark}
\theoremstyle{definition}
\newtheorem{Problem}{Problem}
\theoremstyle{definition}
\newenvironment{parts}{\begin{enumerate}[label=(\alph*)]}{\end{enumerate}}
%tikz	
\tcbset{breakable=true,toprule at break = 0mm,bottomrule at break = 0mm}
\usetikzlibrary{patterns}
\usetikzlibrary{matrix}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
% definitions of number sets
\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\C}{\mathbb{C}}
\allowdisplaybreaks
\begin{document}
	\title{Stochastic Differential Equations}
	\author{Jun Wei Tan}
	\email{jun-wei.tan@stud-mail.uni-wuerzburg.de}
	\affiliation{Julius-Maximilians-Universit\"{a}t W\"{u}rzburg}
	\date{\today}
	\maketitle

\begin{Definition}
	A stochastic differential equation is a (formal) equation of the form
	\begin{equation}\label{eq:SDEdiff}
	\dv{X_t}{t}=b(t, X_t)+\sigma(t, X_t)W_t,
\end{equation}
	where $W_t$ is white noise. 
\end{Definition}
This equation is to be interpreted as follows:
\begin{Definition}
	We say that the stochastic process $X_t$ is a \emph{solution} of the SDE \eqref{eq:SDEdiff} if
	\[X_t = X_0+\int_0^t b(s, X_s)\dd{s}+ \int_0^t \sigma(s, X_s)\dd{B_s}\]
\end{Definition}
\begin{Example}
	A model of a population is given by the stochastic differential equation
	\[\dv{N_t}{t} = r N_t + \alpha W_t N_t.\]
	Here, $r,\alpha$ are constants and $W_t$ is white noise.
\end{Example}
This is the well known model for a population, except that we have allowed $r$ to vary by a white noise term. The solution in the nonstochastic limit is given by a simple exponential. To solve this, we first rewrite the SDE in standard form:
\[\dd{N_t}=r N_t\dd{t}+\alpha N_t \dd{B_t},\]
or
\[\frac{\dd{N_t}}{N_t} = r \dd{t} + \alpha \dd{B_t}.\]
Inspired by the solution in the deterministic case, we guess $g(t, x)=\ln x$ in Ito's formula, and let $Y_t=g(t, N_t)$. Then, we have
\[\dd{Y_t}=\frac{\dd{N_t}}{N_t}+\frac 12 \alpha^2 \dd{t}.\]
Substituting, we have
\[\dd{Y_t} = \left(r- \frac 12\alpha^2 \right)\dd{t} +\alpha B_t\]
which integrates easily to yield
\[N_t = N_0 \exp\left(\left(r - \frac 12 \alpha^2\right)t+\alpha B_t\right).\]
Clearly, for $\alpha=0$, this reduces to the well known exponential solution. This process is known as \uline{geometric Brownian motion}, which, for example, is a solution to the Black-Scholes equation~\cite{inbook}.

Now, we turn to the questions of existence and uniqueness. Recall the basic theorem for existence and uniqueness of a deterministic differential equation
\begin{Theorem}[Picard-Lindelöf]
	Let $\dot{x} = f(t,x)$ be a differential equation with $f$ defined on a rectangle $[a,b]\times \R^n$. If $f$ is Lipschitz continuous in $x$, with Lipschitz constant independent of time, and continuous in time, then the differential equation has a unique global solution on $[a,b]$.
\end{Theorem}
Note that for uniqueness we do not need the continuity in time; the Lipschitz condition alone suffices.
\begin{Definition}[Linear Growth]
	A function $g(t, x)$ on $[a,b]\times \R$ is said to satisfy the \uline{linear growth condition in $x$} if there exists $K>0$ such that
	\[|g(t, x)|\le K(1+|x|)\]
	for all $t\in [a,b]$ and $x\in \R$.
\end{Definition}
Because of the inequality
\[1+x^2 \le (1+x)^2 \le 2(1+x^2),\]
the earlier definition is equivalent to the following definition:
\begin{Proposition}
	A function $g(t,x)$ satisfies the linear growth condition if there exists $K>0$ such that
	\[|g(t,x)|^2 \le K(1+x^2)\]
\end{Proposition}
The second definition will be more useful in the following proofs. Note that it is indeed stronger
\begin{Proposition}
	Globally Lipschitz continuous functions satisfy the linear growth condition.
\end{Proposition}
\begin{proof}
If $f(x)$ is Lipschitz, we have
\[|f(x) - f(0)| \le K |x|,\]
or
\[|f(x)|\le |f(0)| + K|x|\]
by the reverse triangle inequality.
\end{proof}
Conversely, linear growth does not imply Lipschitz continuity, or any kind of continuity
\begin{Example}
	Let
	\[f(x) = \begin{cases}
		x & x\in \Q,\\
		0 & \text{otherwise}.
	\end{cases}\]
	Clearly, $f$ grows linearly, but is not Lipschitz continuous, or even continuous at any point.
\end{Example}

We can prove uniqueness with just Lipschitz continuity, as we do in the deterministic case.
\begin{Theorem}
	Let $\sigma(t,x)$ and $f(t,x)$ be measurable functions on $[a,b]\times \R$ satisfying the Lipschitz condition in $x$. Suppose $\xi$ is an $\mathcal{F}_a$-measurable random variable satisfying $\mathbb{E}[\xi^2]<\infty$. Then the stochastic differential equation
	\[\dv{X_t}{t}=f(t, X_t)+\sigma(t, X_t)W_t\]
	has at most one continuous solution on $[a,b]$.
\end{Theorem}
\begin{proof}
	We will not go through the proof in detail. Instead, we will talk about the main steps. Assume we have two solutions $X_t$ and $Y_t$. We seek to estimate $Z_t=X_t-Y_t$
	\begin{enumerate}
		\item We estimate the expectation value $\mathbb{E}(Z_t^2)$, using the Lipschitz condition.
		\item We obtain an integral inequality
			 \[
			 E(Z_t^2) \le 2K^2 (1+b-a)\int_a^t \mathbb{E}(Z_s^2)\dd{s}
			,\]
			which, by the theory of classical differential equations, implies that $Z_t$ is 0 almost surely for all $t$.
		\item Then, we extend the solution to show that $Z_t$ is $0$ almost surely, using sample path continuity.\qedhere
	\end{enumerate}
\end{proof}
The existence theorem is as follows:
\begin{Theorem}
	The stochastic differential equation \eqref{eq:SDEdiff} has a unique solution with initial condition $\xi$, where $\xi^2$ has finite expectation, $\sigma$ and $b$ are Lipschitz in $x$, with Lipschitz constant independent of $t$, and continuous in $t$.
\end{Theorem}
\begin{proof}
	The proof follows by Picard Iteration much as it does in the deterministic case. We define $X_0=\xi$ and
	\[
	X_t^{(n+1)}=X_0+\int_0^t b(s, X_s^{(n)})\dd{s}+\int_0^t \sigma(s, X_s^{(n)})\dd{B_s}
	.\] 
	Then, the proof proceeds in two steps. First, we show that this sequence converges, If it does, and converges in a dominated manner, it satisfies the integral equation by the dominated convergence theorem. 
\end{proof}
\section{The Markov Property}
It is a known property of initial value problems that the future solution is not dependent on the past. In particular, we can imagine that we have some initial state $x(0)$ and let it evolve a time $t$ to $x(t)$. Then we can let it evolve further. Alternatively, we can consider an initial value problem that has the value $x(t)$ at time $t$. We expect that these two solutions are identical. In a stochastic differential equation, this ```memory'' property is known as the Markov property. Before we define it, let us recall a few definitions:
\begin{Definition}[Conditional Expectation]
	Suppose $(\Omega, \mathcal{F}, \mathbb{P})$ is a probability space and $X$ is a real-valued random variable on that space with finite expectation. Let $\mathcal{H}\subseteq \mathcal{F}$ be a sub $\sigma$-algebra of $\mathcal{F}$. Then the conditional expectation of $X$ relative to $\mathcal{H}$ is a $\mathcal{H}$-$\mathcal{B}$ measurable function $\mathbb{E}(X|\mathcal{H})$ such that
	\[
		\int_H \mathbb{E}(X|\mathcal{H})\dd{\mathbb{P}}=\int_H X \dd{\mathbb{P}}
	\]
	for all $H\in \mathcal{H}$.
\end{Definition}
Examples are as follows
\begin{Example}\noindent

	\begin{center}
		\begin{tikzpicture}
			\begin{axis}[axis lines = left, ytick=\empty, legend pos = south east]
				\addplot[domain=0:2, black, thick]{27*x^3/64-45*x^2/16+9*x/2+2};
				\draw[thick, red] (axis cs:0,{875/256}) -- (axis cs:1,{875/256});
				%\draw[thick, red] (axis cs:1,{965/256}) -- (axis cs:2,{965/256});
				\addplot[domain=1:2, thick, red]{965/256};
				%\draw[thick,blue] (axis cs:0, {2*5947/4096}) -- (axis cs:0.5, {2*5947/4096});
				\addplot[domain=0:0.5, thick, blue]{2*5947/4096};
				\draw[thick, blue] (axis cs:0.5,{2*8053/4096}) -- (axis cs:1, {2*8053/4096});
				\draw[thick, blue] (axis cs:1, {2*8251/4096}) -- (axis cs:1.5, {2*8251/4096});
				\draw[thick, blue] (axis cs:1.5, {2*7189/4096}) -- (axis cs:2, {2*7189/4096});
				\addlegendentry{$X$};
				\addlegendentry{$\mathbb{E}(X|\mathcal{A})$};
				\addlegendentry{$\mathbb{E}(X|\mathcal{B})$};
			\end{axis}
		\end{tikzpicture}
	\end{center}
\end{Example}
Where $Y_1, \dots, Y_n$ is a collection of random variables, we use $\mathbb{E}(X|Y_1, \dots, Y_n)$ to denote $\mathbb{E}(X|\sigma(Y_1,\dots, Y_n))$, the conditional expectation on the $\sigma$-algebra generated by the $Y_i$s.

Using this, we can define the conditional probability of an event,
\begin{Definition}
	The conditional probability of an event $A$, conditional on a $\sigma$-algebra $\mathcal{A}$, $\mathbb{P}(A|\mathcal{A})$ is given by $\mathbb{E}(1_A|\mathcal{A})$.
\end{Definition}
Now, we may define the Markov property
\begin{Definition}
	A stochastic process $X_t$, with $a\le t \le b$, is said to have the \emph{Markov property} if for all sequences $a < t_1 < \dots < t_n < t < b$ and corresponding $x_1, \dots, x_n$, we have
	\[
	\mathbb{P}(X_t\le x|X_{t_1}=x_1, \dots, X_{t_n}=x_n) = \mathbb{P}(X_t\le x|X_{t_n}=x_n)
	.\] 
\end{Definition}
As an example, all processes with independent increments have the Markov property. 
Before we move on, let us prove a useful lemma
\begin{Lemma}
	Suppose a stochastic process $X_t$ defined on $[a,b]$ is adapted to a filtration $\mathcal{F}_t$. Suppose also that
	\[
	\mathbb{P}(X_t\le x|\mathcal{F}_s) = \mathbb{P}(X_t \le x|X_s)
	.\] 
	Then $X_t$ is a Markov process.
\end{Lemma}

Finally, we move to characterise a Markov process by its marginal distributions. Recall that a discrete-time Markov process is characterised by its transition matrix --- that is, the probability to transition to a certain state given its current state.

We define a Markov process on $[a,b]$ and let $a=t_1<t_2$. We have, by Bayes's Theorem
\begin{align*}
	\mathbb{P}(X_{t_1}\le c_1 \cap X_{t_2}\le c_2)=\mathbb{P}(X_{t_2}\le c_2 |X_{t_1}\le c_1)\mathbb{P}(X_{t_1}\le c_1)
\end{align*}
The final property that is of interest to us is time translation invariance. For a deterministic differential equation $\dot{x} = f(x)$, we know that the solution exhibits time translation invariance. In the deterministic case, this is quite easy to see, and follows from the fact that $\dv{t}\varphi(t - t_0) = \varphi'(t-t_0)$. 

In this case, the relevant property is called the \emph{stationary markov property}
\begin{Definition}
	A stochastic process $X$ is called stationary if the moments are time translation invariant:
	\[
	\langle X_{t_1+\tau}X_{t_2+\tau}\dots X_{t_n+\tau}\rangle = \langle X_{t_1}X_{t_2}\dots X_{t_n}\rangle
	\]
	for all $n,\tau$ and $t_1, \dots, t_n$.
\end{Definition}

Thus, we have our final theorem
\begin{Theorem}
	Suppose that $b(x)$ and $\sigma(x)$ are functions satisfying the Lipschitz condition. Then the solution to Eq. \eqref{eq:SDEdiff} is a stationary Markov process.
\end{Theorem}
\bibliographystyle{ieeetr}
\bibliography{ref.bib}
\end{document}

