\begin{Problem}

The BCS wavefunction in momentum space can be written as
\begin{equation}
|\Psi_{\mathrm{BCS}}\rangle
= \prod_{\mathbf{k}} \exp\left( \phi_{\mathbf{k}} c^\dagger_{\mathbf{k}\uparrow} c^\dagger_{-\mathbf{k}\downarrow} \right)|0\rangle
= \prod_{\mathbf{k}} \left(1 + \phi_{\mathbf{k}} c^\dagger_{\mathbf{k}\uparrow} c^\dagger_{-\mathbf{k}\downarrow}\right)|0\rangle .
\end{equation}

\begin{itemize}
\item Show that the BCS ground state is the vacuum for the Bogoliubov quasiparticles, i.e. that the destruction operators $\alpha_{\mathbf{k},\sigma}$ annihilate the BCS ground state. Here
\begin{equation}
\alpha_{\pm\mathbf{k},\sigma} = u_{\mathbf{k}} c_{\pm\mathbf{k}\sigma} + v_{\mathbf{k}} c^\dagger_{\mp\mathbf{k}\overline{\sigma}},
\end{equation}
and you may assume that $u_{\mathbf{k}}, v_{\mathbf{k}}$ are real for convenience.

\item If the Bogoliubov quasiparticle is $\alpha^\dagger_{\mathbf{k}\uparrow} = u_{\mathbf{k}} c^\dagger_{\mathbf{k}\uparrow} + v_{\mathbf{k}} c_{-\mathbf{k}\downarrow}$, then starting from the equation of motion of the Bogoliubov quasiparticle
\begin{equation}
	[H,\alpha^\dagger_{\mathbf{k}\uparrow}] = \pdv{\alpha_{\mathbf{k}\uparrow}^\dagger}{\tau}=E_{\mathbf{k}} \alpha^\dagger_{\mathbf{k}\uparrow},
\end{equation}
show that $(u_{\mathbf{k}}, v_{\mathbf{k}})^T$ is an eigenvector of
\begin{equation}
	h_{\mathbf{k}} = \begin{pmatrix} u_{\mathbf{k}} \\ v_{\mathbf{k}} \end{pmatrix} = \begin{pmatrix} \epsilon_{\mathbf{k}} & \Delta \\ \Delta & -\epsilon_{\mathbf{k}} \end{pmatrix} \begin{pmatrix} u_{\mathbf{k}} \\ v_{\mathbf{k}} \end{pmatrix} = E_{\mathbf{k}} \begin{pmatrix} u_{\mathbf{k}} \\ v_{\mathbf{k}} \end{pmatrix}  
\end{equation}
By solving the eigenvalue problem and assuming the gap is real, give  an explicit expression for $u_{\mathbf{k}}$ and $v_{\mathbf{k}}$ in terms of $\Delta$ and $\epsilon_{\mathbf{k}}$. 
\end{itemize}
\end{Problem}
\begin{proof}
	We apply the annihilation operators directly. For concreteness, we will consider $\alpha_{\va k,\uparrow}$. Since we can commute $\alpha$ through all bilinears $1+\phi_{\va q}c_{\va q}^\dagger c_{\va q}^\dagger$ with $\va q \neq \va k$, we will ignore them all and simply look at the product
	\[( u_{\mathbf{k}} c_{\mathbf{k},\uparrow} + v_{\mathbf{k}} c^\dagger_{-\mathbf{k},\downarrow})(1+\phi_{\va k}c_{\va k,\uparrow}^\dagger c_{-\va k,\downarrow}^\dagger)(1+\phi_{-\va k}c_{-\va k,\uparrow}^\dagger c_{\va k,\downarrow}^\dagger)\]
	applied to the ground state. This product would yield $2^3=8$ terms. However, we can immediately see that some fo them drop out. The product $c_{\va k,\uparrow}$ with the two 1s would yield an annihilation operator applied to the vacuum; it vanishes. We may not multiply $c_{-\va k,\downarrow}^\dagger$ with itself; squares of fermion operators vanish. Thus, two more terms drop out. If we multiply $c_{\va k,\uparrow}$ with something not containing $c_{\va k,\uparrow}^\dagger$, we could simply anticommute it to the end, where it would annihilate the ground state again. Thus, there are only 4 terms left, and we find
	\begin{align*}
		&~( u_{\va{k}} c_{\va{k},\uparrow} + v_{\va{k}} c^\dagger_{-\va{k},\downarrow})(1+\phi_{\va k}c_{\va k,\uparrow}^\dagger c_{-\va k,\downarrow}^\dagger)(1+\phi_{-\va k}c_{-\va k,\uparrow}^\dagger c_{\va k,\downarrow}^\dagger)\\
		=&~u_{\va k}c_{\va k,\uparrow} \phi_{\va k}c_{\va k,\uparrow}^\dagger c_{-\va k,\downarrow}^\dagger + u_{\va k}c_{\va k,\uparrow} \phi_{\va k}c_{\va k,\uparrow}^\dagger c_{-\va k,\downarrow}^\dagger \phi_{-\va k}c^{\dagger}_{-\va k,\uparrow}c_{\va k,\downarrow}^\dagger \\
		&+ v_{\va k}c_{-\va k,\downarrow}^\dagger + v_{\va k}\phi_{-\va k}c_{-\va k,\downarrow}^\dagger \phi_{-\va k}c_{-\va k,\uparrow}^\dagger c_{\va k,\downarrow}^\dagger\\
		=& -u_{\va k}\phi_{\va k}c_{-\va k,\downarrow}^\dagger - u_{\va k}\phi_{\va k}\phi_{-\va k}c_{-\va k,\downarrow}^\dagger c_{-\va k,\uparrow}^\dagger c_{\va k,\downarrow}^\dagger\\
		&+ v_{\va k}c_{-\va k,\downarrow}^\dagger + v_{\va k}\phi_{-\va k}c_{-\va k,\downarrow}^\dagger \phi_{-\va k}c_{-\va k,\uparrow}^\dagger c_{\va k,\downarrow}^\dagger\\
		=&\phi_{-\va k}c_{-\va k\downarrow}^\dagger c_{-\va ,\uparrow}^\dagger c_{\va k,\downarrow}^\dagger (-u_{\va k}\phi_{\va k} +v_{\va k})+c_{-\va k,\downarrow}(-u_{\va k}\phi_{\va k} +v_{\va k})
	\end{align*}
	where we have used the fermionic anticommutation relations to swap $c_{\va k,\uparrow}$ with $c_{\va k,\uparrow}^\dagger$ before pulling it through the rest of the operators to annihilate the ground state. Thus, we see that these operators annihilate the ground state if we have
	\[v_{\va k} = u_{\va k}\phi_{\va k}\]
	In principle, we could use this and the requirement that the transformation stay canonical to determine the parameters $u_{\va k}$ and $v_{\va k}$. But we continue...
	
	Next, we compute the commutator $[H, \alpha_{\va k\uparrow}^\dagger]$ directly. We have previously done a mean field expansion on the BCS Hamiltonian
	\[H = \sum_{\va k,\sigma}\epsilon_{\va k}  c_{\va k,\sigma}^\dagger c_{\va k,\sigma} -\sum_{\va k, \va k'}\frac{g}{V} c_{\va k\uparrow}^\dagger c_{-\va k\downarrow}^\dagger c_{\va k',\uparrow} c_{-\va k', \downarrow} \]
	using the mean field $\Delta = -\frac{g}{V}\langle c_{\va k,\uparrow}c_{-\va k,\downarrow}\rangle$. After decomposing, we find that
	\begin{align*}
		H &= \sum_{\va k,\sigma}\epsilon_{\va k}  c_{\va k,\sigma}^\dagger c_{\va k,\sigma} -\sum_{\va k, \va k'}\frac{g}{V} c_{\va k\uparrow}^\dagger c_{-\va k\downarrow}^\dagger c_{\va k',\uparrow} c_{-\va k', \downarrow} \\
		&= \sum_{\va k,\sigma}\epsilon_{\va k}  c_{\va k,\sigma}^\dagger c_{\va k,\sigma} -\sum_{\va k, \va k'}\frac{V}{g} \left(\overline{\Delta} -\left(\overline{\Delta}-\frac{g}{V}c_{\va k\uparrow}^\dagger c_{-\va k\downarrow}^\dagger\right)\right)\left(\Delta - \left(\Delta - \frac{g}{V}c_{\va k',\uparrow} c_{-\va k', \downarrow}\right)\right)\\
		&= \sum_{\va k,\sigma}\epsilon_{\va k}  c_{\va k,\sigma}^\dagger c_{\va k,\sigma} +\sum_{\va k, \va k'}\left[\overline{\Delta}c_{-\va k \downarrow}c_{\va k\uparrow}+c_{\va k\uparrow}^\dagger c_{-\va k\downarrow}^\dagger \Delta\right]+\frac{V}{g}\overline{\Delta}\Delta
	\end{align*}
In principle, defining the Nambu spinor $\Psi_{\va k} = (c_{\va k \uparrow}, c_{-\va k\downarrow}^\dagger)^T$, we could rewrite the Hamiltonian as (up to constants)
\[H = \begin{pmatrix}
	c_{\va k \uparrow}^\dagger \\ c_{-\va k\downarrow}
\end{pmatrix}\cdot\begin{pmatrix}
	\epsilon_{\va k} & \Delta \\
	\overline{\Delta} & -\epsilon_{\va k}
\end{pmatrix}\begin{pmatrix}
c_{\va k \uparrow} \\ c_{-\va k\downarrow}^\dagger
\end{pmatrix} = \Psi_{\va k}^\dagger \begin{pmatrix}
	\epsilon_{\va k} & \Delta \\
	\overline{\Delta} & -\epsilon_{\va k}
\end{pmatrix}\Psi_{\va k}\]
which would give the desired result. However, we continue as requested: We compute the equation of motion
\[[H, \alpha_{\va k\uparrow}^\dagger]= [H, u_{\va k}c_{\va k\uparrow}^\dagger + v_{\va k}c_{-\va k\downarrow}]=E_{\va k}a_{\va k\uparrow}^\dagger\]
For the kinetic term, we find that
\[[H_\text{kin},  u_{\va k}c_{\va k\uparrow}^\dagger + v_{\va k}c_{-\va k\downarrow}]=\epsilon_{\va k}u_{\va k}c^\dagger_{\va k\uparrow} - \epsilon_{\va k}v_{\va k}c_{-\va k\downarrow}\]
and for the mean field term, we also find that
\begin{align*}
[H_\text{MF},  v_{\va k}c_{-\va k\downarrow}] &= \Delta v_{\va k}[c_{\va k\uparrow}^\dagger c_{-\va k\downarrow}^\dagger, c_{-\va k\downarrow}]\\
&=\Delta v_{\va k}\left(c_{\va k\uparrow}^\dagger c_{-\va k\downarrow}^\dagger c_{-\va k\downarrow}-c_{-\va k\downarrow}c_{\va k\uparrow}^\dagger c_{-\va k\downarrow}^\dagger \right)\\
&=\Delta v_{\va k}\left(c_{\va k\uparrow}^\dagger c_{-\va k\downarrow}^\dagger c_{-\va k\downarrow}+c_{\va k\uparrow}^\dagger (-c^\dagger_{-\va k\downarrow} c_{-\va k\downarrow} + 1)\right)\\
&= \Delta v_{\va k}c_{\va k\uparrow}^\dagger \\
[H_\text{MF},  u_{\va k}c_{\va k\uparrow}^\dagger] &= \overline\Delta u_{\va k}[c_{-\va k\downarrow}c_{\va k\uparrow}, c_{\va k\uparrow}^\dagger]\\
&= \overline\Delta u_{\va k}\left(c_{-\va k\downarrow}c_{\va k\uparrow}c_{\va k\uparrow}^\dagger - c_{\va k\uparrow}^\dagger c_{-\va k\downarrow}c_{\va k\uparrow}\right) \\
&= \overline\Delta u_{\va k}\left(c_{-\va k\downarrow}c_{\va k\uparrow}c_{\va k\uparrow}^\dagger +  c_{-\va k\downarrow}(-c_{\va k\uparrow}c_{\va k\uparrow}^\dagger+1)\right) \\
&=\overline\Delta u_{\va k}c_{-\va k\downarrow}
\end{align*}
Thus, we find that
\[[H, \alpha_{\va k\uparrow}^\dagger] = \epsilon_{\va k}u_{\va k}c^\dagger_{\va k\uparrow} - \epsilon_{\va k}v_{\va k}c_{-\va k\downarrow}+\Delta v_{\va k}c_{\va k\uparrow}^\dagger+\overline\Delta u_{\va k}c_{-\va k\downarrow}\]
By comparison of the above equation with the equation $[H,\alpha_{\va k\uparrow}^\dagger]=E_{\va k}\alpha_{\va k\uparrow}^\dagger$ and comparing the coefficients of $c_{\va k\uparrow}^\dagger$ and $c_{\va k\uparrow}$, we find that $(u_{\va k}, v_{\va k})$ is an eigenvector of the desired matrix.
\end{proof}
\begin{Problem}

One useful way to regard superconductors is via the semiconductor analogy, in which the quasiparticles are treated like the positive and negative energy excitations of a semiconductor. To do so we divide the Brillouin zone up into equal halves and redefine a set of positive and negative energy quasiparticle operators according to
$\alpha^\dagger_{\mathbf{k},\sigma+} = \alpha^\dagger_{\mathbf{k}\sigma}$ and $\alpha^\dagger_{\mathbf{k}\sigma-} = \operatorname{sgn}(\sigma)\,a_{-\mathbf{k}-\sigma}
\text{ for } \mathbf{k} \in \tfrac12 \mathrm{BZ}.$

\begin{itemize}
\item By this procedure rewrite the BCS Hamiltonian in terms of these new operators, and show that the excitation spectrum can be interpreted in terms of an empty band of positive energy excitation and a filled band of negative energy excitations.

\item Show that the BCS ground state wave functions can be regarded as a filled sea of negative energy quasiparticle states and an empty sea of positive energy quasiparticle states.
\end{itemize}
\end{Problem}
\begin{proof}
For reference the BCS Hamiltonian referred to in the problem is the already diagonalised BCS mean field Hamiltonian in the quasiparticle basis
\[H = \sum_{\va k}E_{\va k}(a_{\va k\uparrow}^\dagger a_{\va k\uparrow} - a_{-\va k\downarrow}a_{-\va k\downarrow}^\dagger)+V\frac{\overline{\Delta}\Delta}{g_0}\]
\end{proof}

\begin{Problem}

\begin{itemize}
\item Assuming that the Debye frequency is a small fraction of the bandwidth, show that the difference between the superconducting and normal-state free energies can be written as
\begin{equation}
\mathcal{F}_S - \mathcal{F}_N =
-2T N(0) \int_{-\omega_D}^{\omega_D}
\ln\!\left[
\frac{\cosh\!\left(\sqrt{\varepsilon^2+|\Delta|^2}/2T\right)}
{\cosh(\varepsilon/2T)}
\right] d\varepsilon
+ V \frac{|\Delta|^2}{g_0}.
\end{equation}

Notice that the previous expression is invariant under changes in the phase of the gap parameter. What consequences does it imply?

\item By differentiating the expression of the Free energy with respect to $\Delta$, confirm the zero temperature gap equation
\begin{equation}
\frac{V}{g_0 N(0)} =
\int_0^{\omega_D} \frac{d\varepsilon}{\sqrt{\varepsilon^2+\Delta_0^2}},
\end{equation}
where $\Delta_0 = \Delta(T = 0)$. Use this result to eliminate $g_0$, to show that the free energy can be written as
\begin{equation}
\mathcal{F}_S - \mathcal{F}_N =
N(0)\Delta_0^2\,
\Phi\!\left[\frac{\Delta}{\Delta_0},\frac{T}{\Delta_0}\right],
\end{equation}
where the dimensionless function
\begin{equation}
\Phi[\delta,t] =
\int_0^\infty dx
\left\{
-4t \ln
\frac{\cosh\!\left(\sqrt{x^2+\delta^2}/2t\right)}
{\cosh(x/2t)}
+ \frac{\delta^2}{\sqrt{x^2+1}}
\right\}.
\end{equation}

What implies the change in the integration boundary to infinity?
\end{itemize}
\end{Problem}
\begin{proof}
	The expression for the free energy is assumed to be known in the question: It can be derived by performing a Hubbard-Stratonovich transformation, integrating out the fermions and then performing a mean field substitution on the Hubbard-Stratonovich field, with which the path integral disappears and only the product of eigenvalues forming a determinant remain.
	
	The free energy is given by
	\[\mathcal{F}[\Delta, T] = -T\sum_{\va k, n}\ln[\omega_n^2+\epsilon_{\va k}^2+|\Delta|^2]+V\frac{|\Delta|^2}{g_0}\]
	where it is to be understood that the momentum sum is restricted to a shell around the Fermi surface - while the free energy would involve all electrons, not just those in this small shell, it makes no difference when finding minima, as they are noninteracting and hence behave like free electrons. In the path integral, this would be understood as simply integrating out these fermions and ignoring the constants resulting therefrom. Carrying out the Matsubara sum, it can be found that (as I understand, this result is also assumed)
	\[\mathcal{F}= -2TV\int_{|\epsilon_{\va k}|<\omega_D}\frac{\dd[3]{k}}{(2\pi)^3}\left[\ln[2\cosh(\beta E_{\va k}/2)]\right]+V\frac{|\Delta|^2}{g_0}\]
	where $E_{\va k}$ is the quasiparticle dispersion $E_{\va k} = \sqrt{\epsilon_{\va k}^2 + \Delta^2}$ (assuming the gap parameter $\Delta$ is real).
	
	The normal state free energy is just the free energy evaluated at zero gap parameter $\Delta$. Substituting, we easily find that
	\[\mathcal{F}_S-\mathcal{F}_N = -2TV \int \frac{\dd[3]{k}}{(2\pi)^3}\ln \left[\frac{\cosh\left(\sqrt{\varepsilon^2+|\Delta|^2}/2T\right)}
	{\cosh(\varepsilon/2T)}\right]+V\frac{|\Delta|^2}{g_0}\]
	The last step in this derivation is to replace the integration measure with an integral over the energy and the density of states $V\frac{\dd[3]{k}}{(2\pi)^3}=N(\epsilon)\dd{\epsilon}$ and approximate the density of states as constant within a tiny shell around the Fermi surface to get
\[\mathcal{F}_S - \mathcal{F}_N =
-2T N(0) \int_{-\omega_D}^{\omega_D}
\ln\!\left[
\frac{\cosh\!\left(\sqrt{\varepsilon^2+|\Delta|^2}/2T\right)}
{\cosh(\varepsilon/2T)}
\right] d\varepsilon
+ V \frac{|\Delta|^2}{g_0}.\]
	where $N(0)$ is the density of states at the Fermi surface.
	
	Next, we differentiate the free energy as requested to find
	\begin{align*}
		\dv{\Delta}\Delta\mathcal{F}=&~ -2TN(0)\int_{-\omega_D}^{\omega_D} \tanh\left(\sqrt{\varepsilon^2+|\Delta|^2}/2T\right)\frac{\Delta}{2T\sqrt{\epsilon^2+\Delta^2}}\\
		&~+\frac{2V\Delta}{g_0}=0
	\end{align*}
Noting that $\tanh(1/T)\to 1$ as $T\to 0$, we have
\[N(0)\int_{-\omega_D}^{\omega_D} \frac{1}{\sqrt{\epsilon^2+\Delta^2_0}}\dd{\epsilon}=\frac{2V}{g_0}\]
Using the evenness of the integrand to allow us to integrate over only half of the specified interval, this reduces to
\[\frac{V}{g_0 N(0)} =
\int_0^{\omega_D} \frac{\dd\varepsilon}{\sqrt{\varepsilon^2+\Delta_0^2}}\]
as desired.

The final expression we are asked to show follows by simply  substituting to find
\begin{align*}
	\mathcal{F}_S - \mathcal{F}_N &= -2T N(0) \int_{-\omega_D}^{\omega_D}
	\ln\!\left[
	\frac{\cosh\!\left(\sqrt{\varepsilon^2+|\Delta|^2}/2T\right)}
	{\cosh(\varepsilon/2T)}
	\right] \dd\varepsilon
	+ |\Delta|^2N(0)\int_0^{\omega_D} \frac{\dd\varepsilon}{\sqrt{\varepsilon^2+\Delta_0^2}}\\
	&=N(0)\int_0^{\omega_D} -4T\ln\!\left[
	\frac{\cosh\!\left(\sqrt{\varepsilon^2+|\Delta|^2}/2T\right)}
	{\cosh(\varepsilon/2T)}\right] + \frac{|\Delta|^2}{\sqrt{\epsilon^2+\Delta_0^2}}\dd{\epsilon}
\end{align*}
The derivation of the desired expression
\[\mathcal{F}_S - \mathcal{F}_N =
N(0)\Delta_0^2\,
\Phi\!\left[\frac{\Delta}{\Delta_0},\frac{T}{\Delta_0}\right]\]
is now obvious. We note that during this derivation, we would replace $\epsilon/ \Delta_0$ with $x$; in the case where the gap parameter is small, this would extend the integration bounds to infinity.
\end{proof}

\begin{Problem}

Consider two superconductors which are connected via a small bridge, also superconducting, of length $L$. Each of those superconductors can be thought as a lead with all of its properties well defined and each of the wave function in this bridge
\begin{equation}
\Psi(x \to 0) = \Psi_1 = |\Psi_0|e^{i\phi_1}
\end{equation}
\begin{equation}
\Psi(x \to L) = \Psi_2 = |\Psi_0|e^{i\phi_2}
\end{equation}
for the edges. In this area the wave function can be thought as a superposition of those two
\begin{equation}
\Psi(x) = |\Psi_0|e^{i\phi_1} g(x) + |\Psi_0|e^{i\phi_2}(1-g(x)),
\end{equation}
with the only requirement that $g(x \to 0) \to 1$ and $g(x \to L) \to 0$. We aim to get the behaviour of $g(x)$, in order to do so use the Ginzburg Landau equation
\begin{equation}
-\frac{\hbar^2}{2m}\frac{\partial^2\Psi}{\partial x^2}
+ \alpha \Psi + \beta|\Psi|^2\Psi = 0,
\end{equation}
to have a complete description.

\textit{Hint:} It might be useful to think of $\Psi(x)=|\Psi_0|f(x)$ and rewrite an equation for $f(x)$.
Use $|\Psi_0|^2=-\alpha/\beta$ and define also $\xi^2=\hbar^2/(2m|\alpha|)$. Use the hypothesis of short link $L \ll \xi$.
\end{Problem}
\begin{proof}
	As suggested, we think of $\Psi(x)$ as $|\Psi_0|f(x)$. Then, we substitute this into the Ginzburg-Landau equation to find
	\begin{align*}	
		0&=-\frac{\hbar^2}{2m}\pdv[2]{f(x)}{x}+ \alpha f(x) + \beta |\Psi_0|^2 |f(x)|^2f(x)\\
		&=-\frac{\hbar^2}{2m}\pdv[2]{f(x)}{x}+ \alpha f(x) -\alpha |f(x)|^2f(x)\\
		&= \zeta^2 \pdv[2]{f(x)}{x}+f(x)(1-|f(x)|^2)
	\end{align*}
The first step is to de-dimensionalise this equation. We do so by substituting $x \to Ly$, which yields $\dv{x} = \frac 1L \dv{y}$. Substituting, we find that
\[\pdv[2]{f(y)}{y} + \epsilon^2 f(y)(1-|f(y)|^2)=0\]
where $\epsilon = L / \zeta$ is the small parameter. Then, we substitute a WKB style ansatz
\[f(y) = A(y) e^{S(y)}\]
to find

\end{proof}

\begin{Problem}
With the previous result, compute the Josephson current. Show that it depends exclusively on $\phi =\phi_2-\phi_1$ and not on each phase individually. What is the expression for the free energy?	
\end{Problem}

\begin{Problem}
	How is the current changed when one applies a voltage? Consider the Schr\"{o}dinger equation and assume $\Psi(x,t) = |\Psi(x)|e^{i\phi(t)} $. Use that to find $\phi(t)$ assuming $\phi(t=0)=\phi_0$. You may associate the energy of the external superconductors with their chemical potential, and the difference of chemical potential with the applied voltage.	
\end{Problem}
